---
title: "Experiment to study basics of R Programming."
output: html_notebook
author: "Harshita Singh"
date: "25-03-2021"
---

**R Basics :** Comments, variables, identifier, keywords, basic data types, type conversion, I/P & O/P statements, formatting output etc.

**Programs :**

1. Program to calculate simple and compound interest.
2. Program to performs different conversions like temperature conversion, distance conversion etc.

```{r}
# This is a comment
"Hello World"

"Single Line Cooment" #This is a comment

# Multiline Comments
# are written
# this way in R
"Multiline Comments"
```
```{r}
# Creating a variable in R
name <- "John Doe"
age <- 21

# Print/Output Variables
name
print(age)
```
```{r}
# Concatenate Elements
text <- "easy"
paste("R is ", text)

text1 <- "Learning R is"
paste(text1, text)

```
```{r}
# Multiple Variables
var1 <- var2 <- var3 <- "Assignment"
var1
var2
var3
```

```{r}
# Legal variable names:
myvar <- "John"
my_var <- "John"
myVar <- "John"
MYVAR <- "John"
myvar2 <- "John"
.myvar <- "John"

```

```{r}
# Basic Data Types
# numeric
x <- 10.5
class(x)

# integer
x <- 1000L
class(x)

# complex
x <- 9i + 3
class(x)

# character/string
x <- "R is exciting"
class(x)

# logical/boolean
x <- TRUE
class(x)

```

Program to calculate simple and compound interest.

**Note :** Run the program on the console.
```{r}
while (TRUE) {
  # Printing the menu
  cat("Menu : \n1. Calculate simple interest \n2. Calculate compound interest \n0. Exit \nEnter your choice : ")
  choice <- scan()
  
  if (choice == 0)  {
    break 
  } else if (choice == 1 || choice == 2) {
    # Calculate SI or CI
    # Formula :  SI = p*n*r/100, CI = p*((1+r/100)^n)
    print("Enter principal amount : ")
    p <- scan()
    
    print("Enter rate : ")
    r <- scan()
    
    print("Enter time span : ")
    n <- scan()
    
    if (choice == 1) {
      cat ("Simple Interest = ", p*n*r/100, "\n\n")
    } else if (choice == 2) {
      cat("Compound Interest = ", p*((1+(r/100))^n), "\n\n")
    }
    
  } else {
    print("Wrong Choice!")
    break
  }
}

```

Program to perform different conversions like temperature conversion, distance conversion etc.

**Note :** Run the program on the console.
```{r}
while (TRUE) {
  # Printing the menu
  cat(
    "Menu: 
    \n1. Convert celsius to fahrenheit 
    \n2. Convert fahrenheit to celsius 
    \n3. Convert from Feet to Meters
    \n4. Convert from Feet to Inches
    \n5. Convert from Meters to Feet
    \n6. Convert from Meters to Inches
    \n0. Exit
    \nEnter your choice : "
  )
  
  choice <- scan()
  
  if (choice == 0 || length(choice) != 1) {
    
    print("Exiting...")
    break
    
  } else if (choice == 1) {
    
    # Convert celsius to fahrenheit 
    print("Enter temperature in Celsius : ")
    t <- scan()
    
    cat("Temperature in fahrenheit = ", ((9*t) + (32*5))/5, "\n")
    
  } else if (choice == 2) {
    
    # Convert fahrenheit to celsius 
    print("Enter temperature in fahrenheit : ")
    t <- scan()
    
    cat("Temperature in celsius = ", (5*(t - 32)/9), "\n")
    
  } else if (choice == 3) {
    
    # Convert from Feet to Meters 
    print("Enter distance in feet : ")
    t <- scan()
    
    cat("Distance in meters = ", t/3.2808, "\n")
    
  } else if (choice == 4) {
    
    # Convert from Feet to Inches 
    print("Enter length in feet : ")
    t <- scan()
    
    cat("Length in inches = ", t*12, "\n")
    
  } else if (choice == 5) {
    
    # Convert from Meters to Feet 
    print("Enter distance in meters : ")
    t <- scan()
    
    cat("Distance in feet = ", t*3.2808, "\n")
    
  } else if (choice == 6) {
    
    # Convert from Meters to Inches 
    print("Enter length in Meters : ")
    t <- scan()
    
    cat("Length in inches = ", t*39.370, "\n")
    
  } else {
    cat("Wrong Choice! \n")
  }
}
```



